---
output:
  html_document: default
  pdf_document: default
---
# Homework 4
**UNI:** qy2205  
**Name:** Quan Yuan  
**Email:** qy2205@columbia.edu  


```{r}
debt <- read.csv("debt.csv", as.is = TRUE)
dim(debt)
head(debt)
```
### 1
```{r}
# (a)
library(plyr)
mean.growth <- function (data){
  return(mean(data$growth))
}
# (b)
country.ave.growth <- signif(daply(debt, .(Country), mean.growth), 3)
country.ave.growth
country.ave.growth[c("Australia","Netherlands")]
```
#### Report: the average GDP growth rates is highest in Japan and lowest in UK. Austria, Jpan and Portugal have high average GPA growth rate (above 4). UK, Greece and Denmark have low GPA growth rate (below 3).

### 2
```{r}
each.year.ave.growth <- signif(daply(debt, .(Year), mean.growth), 3)
each.year.ave.growth
each.year.ave.growth[c("1972","1989")]
```
```{r}
year <- as.numeric(names(each.year.ave.growth))
year.ave.gdp <- unname(each.year.ave.growth)
plot(year, year.ave.gdp, xlab = 'Year', ylab = 'Average GDP', main = 'the average GDP growth rate for each year')
```
### 3
```{r}
# (a)
cat('the correlation coeffcient between GDP growth and the debt ratio over the whole data set', signif(cor(debt$growth, debt$ratio), 4))
# (b)
debt.gdp.cor <- function (data) {
  return(cor(data$growth, data$ratio))
}
cor.country <- daply(debt, .(Country), debt.gdp.cor)
signif(cor.country, 3)
cat('The mean of correlations (country):', signif(mean(cor.country),4), '3 digits:', signif(mean(cor.country),3))
hist(unname(cor.country), breaks = 10, xlab = 'correlations')
# (c)
cor.year <- daply(debt, .(Year), debt.gdp.cor)
signif(cor.year, 3)
cat('The mean of correlations (year):', signif(mean(cor.year),4), '3 digits:', signif(mean(cor.year),3))
hist(unname(cor.year), breaks = 10, xlab = 'correlations')
# (d)
cat('Country: Norway has very high positive corr, Jpan and Germany have high negative corr')
cat('Year: 1946 and 1957 have high negative corr, 1978 has high positive corr')
```
### 4
```{r}
lm0 <- lm(debt$growth ~ debt$ratio, data = debt)
cat('slope is ', signif(lm0$coefficients[2], 3))
cat('intercept is ', signif(lm0$coefficients[1], 3))
plot(debt$ratio, debt$growth, xlab = 'ratio', ylab = 'growth', main = 'growth rate against debt ratio')
abline(lm0, col = 'red')
```
### 5
```{r}
four.countries <- names(cor.country[cor.country < -0.5])
country.plot <- function (data){
  plot(data$growth, data$ratio, main = unique(data$Country), xlab = 'growth', ylab = 'ratio')
  return(NA)
}
par(mfrow = c(2,2))
daply(debt[debt$Country %in% four.countries, ], .(Country), country.plot)
```
### 6
```{r}
# (a)
France <- debt[debt$Country == 'France', ]
dim(France)
# (b)
France$judge <- ifelse((c(France$Year[2:nrow(France)] - 1, NA) == c(France$Year)) == TRUE, 0, NA)
France$next.growth <- c(France$growth[2:nrow(France)], NA) + France$judge
France <- subset(France, select = -judge)
cat('Year 1971:', signif(France[France$Year == 1971,]$next.growth, 3))
cat('Year 1972:', France[France$Year == 1972,]$next.growth)
```
### 7
```{r}
get.next.growth <- function (data) {
  data$judge <- ifelse((c(data$Year[2:nrow(data)] - 1, NA) == c(data$Year)) == TRUE, 0, NA)
  data$next.growth <- c(data$growth[2:nrow(data)], NA) + data$judge
  data <- subset(data, select = -judge)
  return(data)
}
data.list <- ddply(debt, .(Country), get.next.growth)
cat('France Year 2009:', data.list[(data.list$Country == 'France')&(data.list$Year == 2009), ]$next.growth)
```
### 8
```{r}
plot(data.list$ratio, data.list$next.growth, xlab = 'ratio', ylab = 'next year growth', main = 'ratio ~ next year growth')
lm1 <- lm(data.list$next.growth ~ data.list$ratio)
abline(lm1, col = 'red')
cat('Intercept is:',signif(lm1$coefficients[1], 3))
cat('Slope is:',signif(lm1$coefficients[2], 3))
```
#### the slope becomes closer to zero, which means the the next year GDP growth is less related to the current debt ratio. The intercept becomes samller, which means the next year GDP growth is likely to be lower than this year if debt ratio is 0

### 9
```{r}
plot(data.list$growth, data.list$next.growth, xlab = 'this year growth', ylab = 'next year growth', main = 'this year growth ~ next year growth')
lm2 <- lm(data.list$next.growth ~ data.list$growth)
abline(lm2, col = 'red')
```
#### Answer: current growth is a better predictor of future growth since it has a more significant slope.



