---
title: "HW3"
author: "Zeyu Yang, zy2327"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## i Save source code into a html file
```{r}
# setwd("/Users/zeyu/Google Drive/School/Statistical Computing & Introduction to DataScience GR5206_003/Homeworks/HW3")
```
The source code has saved as _NetsSchedule1819.html_ in the above folder. 
And it is checked that it can be opened in text editor.

## ii
```{r}
# Read Data
nets1819 <- readLines('NetsSchedule1819.html', warn = FALSE)
```

```{r}
#a
l <- length(nets1819)
cat("There are",l,"lines.")
#b
n <- sum(nchar(nets1819))
cat("There are",n,"characters.")
#c
m <- max(nchar(nets1819))
cat("The maximum number of characters in a single line of the file is",m,".")
```

## iii
They will play with Detriot first at 7:00PM Wed, Oct 17.

They will play with Miami last at 8:00PM Wed, Apr 10.

## iv
_Line 64_ in the file holds information about the game of the regular season.
This line starts with
```
                <div id="espnfitt"><div class="bp-mobileMDPlus bp-mobileLGPlus bp-tabletPlus bp-desktopPlus bp-desktopLGPlus" data-fitt-page-type="teamSchedule" data-reactroot=""><div id="espn-analytics">
```

Ends with:
```
<div id="fittOverlayContainerInner" data-box-type="fitt-adbox-overlay"><div class="Ad ad-slot" data-slot-type="overlay" data-out-of-page="true" data-slot-kvps="pos=outofpage"></div></div></div></div></div></div></div></div>
```
## v
```{r}
# Use the first game date to extract the line
line_pattern <- "Wed, Oct 17"
#Extract the line
line <- nets1819[grep(line_pattern,nets1819)]
```

## vi
```{r}
# Around the end of each row, there will be a "Schedule__ticket". So we can use these characters to split the line
line1 <- strsplit(line, split = "Schedule__ticket")
# Since there are some redundant codes in the last part, we may remove it.
line2 <- line1[[1]][1:(length(line1[[1]])-1)]
cat("The line had split into",length(line2),"rows.")
```
```{r}
#Alternative Method
#pattern <- "<tr class=\"Table2__tr Table2__tr--sm Table2__even\" data-idx=\"[0-9]{1,2}\">.+?</a></td></tr>"
#line1 <-  regmatches(line,gregexpr(pattern,line)) 
```

## vii
```{r}
# Create pattern
date_pattern <- "[A-Z][a-z][a-z],\\s[A-Z][a-z][a-z]\\s[0-9]{1,2}"
#Find lines
day <- grep(date_pattern, line2)
cat("There is",length(day),"lines.")
# Match first game
cat("Match the first game:", regmatches(line2[1],regexpr(date_pattern,line2[day[1]])) == "Wed, Oct 17")
# Match last game
cat("Macth the last game:",regmatches(line2[82],regexpr(date_pattern,line2[day[82]])) == "Wed, Apr 10")
```

## viii
```{r}
#Extract the dates and store it into vector date
date <- regmatches(line2[day],regexpr(date_pattern,line2[day]))
date
```

## ix Store the time of game
```{r}
# Create pattern
time_pattern <- "[0-9]{1,2}:[0-9][0-9]\\s[A|P]M"
#Extract the time and store it into vector time
time <- regmatches(line2,regexpr(time_pattern,line2))
time
```

## x Home/Away
```{r}
pattern <- "<span class=\"pr2\">(@|vs)"
#Extract information
home1 <- regmatches(line2,regexpr(pattern,line2))
home <- ifelse(home1=="<span class=\"pr2\">vs",1,0)
home
```

## xi Opponent
```{r}
pattern <- "\">.{1,20}<!-- --> </a></span></div></td>"
temp <- regmatches(line2,regexpr(pattern, line2))
temp1 <- strsplit(temp, split = ">|<")
opponent <- c()
for(i in 1:length(temp1))
{
  opponent <- c(opponent,temp1[[i]][2])
}
opponent
```

## xii
```{r}
Schedule <- data.frame(date,time,opponent,home)
head(Schedule,10)
```

Yes, it matches the first 10 rows.